\chapter*{Acknowledgments}

There is an old saying among parents: It takes a village to raise a child. The
same phrase likely applies to the development of any dissertation, and as such
the author is grateful to his fellow villagers listed here.

First, the author's committee requires special acknowledgment and gratitude
for taking on an nontraditional student and an interdisciplinary topic.
Neither are easy and they did both at the same time. The committee includes Jack
Dongarra, John Drake, Mike Guidry (chair), Mallikarjun Shankar, and John Turner. 

The author is grateful for the support provided by the US Department
of Energy and the ORNL Director’s Research and Development Fund in the
Integrated Computational Environment for the Modeling and Analysis of Neutrons
(ICEMAN) project. ORNL is managed by UT-Battelle LLC for the US Department of
Energy under contract DE-AC05-00OR22725. The completion of this work would not
have been possible without the continued encouragement and uplifting support of
colleagues and management at ORNL. The author is also grateful to the management
and staff of RNET Technologies LLC, especially Gerald Sabin and Ben O'Neill, who
supported part of this work under subcontract to ORNL.

The author would like to express his sincere thanks to his coauthors and
fellow developers of the Eclipse Integrated Computational Environment that
forms much of the foundation of this work. These collaborators include Andrew
Bennett, Jordan Deyton, Kasper Gammeltoft, Jonah, Graham, Dasha Gorin, Hari
Krishnan, Menghan Li, Alexander J. McCaskey, Taylor Patterson, Robert Smith,
Gregory Watson, and Anna Wojtowicz. Other collaborators who the author would
like to acknowledge for thoughtful discussions on workflows include Jim Belak on
the nature of workflows in the ExAM project, and Robert Clay, Dan Laney, and
David Montoya on modeling and simulation workflows. The earliest work on this
project included substantial effort by many others who directly or
indirectly contributed to the project, either in its early days as “NiCE” or
once it moved to Eclipse. This includes Ronald Allen, Andrew Belt, David E.
Bernholdt, Tim Bohn, Erica Grant, John M. Hetrick III, Forest Hull, Sebastien
Jourdain, JiSoo Kim, Allison Koenecke, Fangzhou Lin, Eric J. Lingerfelt, Greg
Lyon, Tony McCrary, Elizabeth Piersall, Neeti Pokhriyal, Adrian Sanchez, Claire
Saunders, Nick Stanish, Matthew Wang, ivand Scott Wittenberg. The authors would
like to acknowledge the special contribution of the Eclipse Foundation, the
Eclipse Community, the Eclipse Science Working Group, and our many colleagues
who use and contribute to open-source projects in the Eclipse ecosystem.
Finally, the development team is especially grateful to Barney Maccabe, David
Pointer for their endless support and advocacy for this work.

The author is grateful for the information and citations on Scrybe from Anthony
Skjellum, Director of the SimCenter at the University of Tennessee,
Chattanooga. The author is also grateful to the internal reviewers and helpful
colleagues at the Oak Ridge National Laboratory, and many other anonymous
reviewers who provided feedback on the journal articles and conference
presentations that sprung from this work.

Shantenu Jha of Rutgers and Brookhaven National Laboratory, and Katie Knight, of
Oak Ridge National Laboratory, deserve their own acknowledgements chapter and
together could easily serve as the model for an entire ontology for ``Ideal
Colleagues.'' The kindness, encouragment, and continued support, as well as the
constant willingness to act as a sounding board for crazy ideas has earned these
perpetual thanks and beer at the authors expense.

Charlie Horak edited this work, including the author's bad references to Star
Trek in Chapter \ref{ch:ontologies}, and the author is grateful for her patience
dealing with inconsistent commas, tense switching, bad numbering, and a number
of other things that are not in the version the reader is holding thanks to her
efforts.

An extremely large amount of peanut butter and chocolate was consumed in the
creation of this work. This includes a very large amount of the Peanut Butter
Fudge available at the Island Scoop Restaurant in Jamestown Rhode Island. The
author appreciates the kindness of the proprietor(s) and staff who let him crash
at the restaurant for several hours a day while he worked on corrections during
his family vacation.

Last but not least, the author would like to acknowledge the sacrifies made by
and continuous support and love from his wife and daughter, both of whom
constantly encouraged him, even on vacation, to complete this work. Of all the
villagers who made this possible, none did more than these two and the author
expresses his extreme gratitude and love to them.
