\chapter*{Vita} \label{ch:vita}

Jay Jay Billings is a Research Scientist in the Neutron Scattering
Division (NSD) and Computer Science and Mathematics Division (CSMD) at Oak Ridge
National Laboratory. He leads the Scientific Computing and Software Engineering group
in NSD and Research Software Engineering group in CSMD. He holds the Bachelor’s
Degree in Physics from Virginia Tech, class of 2005, and the Master of Science
in Theoretical Astrophysics from the University of Tennessee, class of 2008.
Mr. Billings’ research focuses on the design and implementation of modeling and
simulation tools for energy science, a large part of which has been related to
the study of scientific workflows in an HPC context.

At Oak Ridge National Laboratory, Mr. Billings is leading the Scientific
Software Initiative within CSMD and is a member of the Software Council, both of
which are taking a new look at the way ORNL develops software for the Department of
Energy. He is a founding member and current chair of the Science Working Group
at the Eclipse Foundation, where he also leads the Eclipse Integrated
Computational Environment and the Eclipse Advanced Visualization Project. Mr.
Billings was also appointed to the Eclipse Architecture Council in 2016, and is
a mentor for several additional Eclipse projects. He is a member of the
Association for Computing Machinery.

Mr. Billings has been funded by the Department of Energy Offices of Nuclear
Energy, Energy Efficiency and Renewable Energy, Advanced Scientific Computing
Research, Basic Energy Sciences, and Advanced Manufacturing.

In addition to his day job, Mr. Billings is a candidate for the PhD in Energy
Science from the Bredesen Center for Interdisciplinary Research and Education at
the University of Tennessee. He spends his spare time with his family, and
singing.
